---
title: "rmd_live_code"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Live Code!

This is live! Meaning anytime we save this file, it will appear online in a plain text format, [**here**](https://dl.dropboxusercontent.com/s/fbcybzmhb7phgph/r4wrds_live_rmd.Rmd?dl=0).

Notes to ourselves about what we are doing. **Bold**, *italic*

```{r libraries}

# code as we would normally do in R script
library(here)
library(sf)
library(dplyr)
library(readr)
library(viridis)
library(ggplot2)
library(USAboundaries)
library(rnaturalearth)
library(GSODR)
library(cowplot)
library(mapview)      # interactive maps!
mapviewOptions(fgb = FALSE)

```

Now let's load some data. 

```{r}

# get USA states, filter out Puerto Rico, Alaska, and Hawaii for now
us <- USAboundaries::us_boundaries(type = "state", resolution = "low") %>% 
  filter(!state_abbr %in% c("PR", "AK", "HI"))

# get CA boundary with high definition
ca <- USAboundaries::us_states(resolution = "high", states = "CA")

# make a box around CA (a grid with an n=1) for inset
ca_box <- st_make_grid(ca, n = 1)

# get crs
st_crs(ca_box)$epsg

# get CA county boundary
ca_co <- USAboundaries::us_counties(resolution = "high", states = "CA")

# mapview(ca_box) # quick check that it worked
# mapview of two obj
# mapview(ca_box) + mapview(ca)

# make sure we have all the pieces with a quick test plot
# check they all have same crs:
library(purrr)
map(list(us, ca, ca_co), ~st_crs(.x)$epsg)

# make a map
plot(us$geometry) # every time this will make a fresh map
plot(ca$geometry, add = TRUE, col = "gray50", border = "maroon")
plot(ca_co$geometry, add = TRUE, border = "pink", col = NA)
plot(ca_box, add = TRUE, border = "red3", col = NA, lwd = 2)


```

## Reading in Spatial Data in R

```{r}

# can do this iteratively
sac_co <- st_read(here("data/shp/sac/sac_county.shp"))

mapview(sac_co)

```

What about with geopackages?
We can look and see what layers exist in these spatial databases with `st_layers`

```{r gpkg}

gpkg <- here("data/sacramento_gsod.gpkg")

st_layers(gpkg)

gpkg_sj <- here("data/san_joaquin_gsod.gpkg")
st_layers(gpkg_sj)

```

## Visualize


```{r}

climdata_sac <- st_read(dsn = gpkg, 
                        layer = "sacramento_isd_data")

climdata_isd_pts <- st_read(dsn = gpkg, 
                            layer = "sacramento_isd_pts")

mapview(climdata_isd_pts)

```

Now we can make a plot data tied to a spatial point:

```{r}
# MONTHLY AVG:
clim_month <- climdata_sac %>% 
  filter(!is.na(PRCP)) %>% # filter out missing data
  filter(!is.na(TEMP)) %>% 
  group_by(NAME, STNID, MONTH) %>% 
  summarize(
    across(c("TEMP","PRCP"), list(min = min, mean = mean, max = max))
  ) %>% 
  ungroup()

# monthly prcp
(mPPT <- ggplot() + 
  geom_col(data = clim_month, 
           aes(x = MONTH, y = PRCP_mean,
               fill = PRCP_mean), show.legend = T) +
  theme_minimal() + 
  labs(y = "", x = "") +
  scale_x_continuous(breaks = c(1, 4, 7, 10),
                     labels = c("Jan", "Apr", "Jul", "Oct")) +
  theme(plot.background = element_blank(),
        legend.position = "left",
        #legend.position = c(-0.25, 0.55),
        legend.key.height = unit(.15, units = "in"),
        legend.key.width = unit(.1, units = "in"), 
        panel.border = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = unit(c(0, 0, 0 ,0), "mm")) +
  scale_fill_viridis_c("Mean \nPPT(in)") +
  labs(subtitle = "Monthly Mean Precipitation (1960-2020)",
       caption = 
         "Data from the {GSODR} package: https://ropensci.github.io/GSODR/")+
  coord_polar() + 
  facet_wrap(clim_month$NAME~., nrow = 1))


```


# Spatial Joins

```{r}

# groundwater level data from stations in sacramento
gwl_sac <- read_csv("data/gwl/gwl_sac.csv")

# make spatial
gwl_sac <- gwl_sac %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), 
           crs = 4326, remove = FALSE)  

mapview(gwl_sac) # this may take a minute because of duplication

# make a distinct stations dataset
gwl_sac_stations <- gwl_sac %>% 
  select(SITE_CODE, WLM_GSE, contains("DESC"), WELL_NAME:COUNTY_NAME) %>% 
  distinct(.keep_all = TRUE)

mapview(gwl_sac_stations, zcol="WLM_DESC")
st_crs(gwl_sac_stations)$epsg
```

## Bring Census Tract data for CalEnviroscreen

```{r}

ces <- st_read(here("data/calenviroscreen/CES3_shp/CES3June2018Update.shp"))
st_crs(ces)$epsg # diff proj
mapview(ces)


```

## Crop The Data to Sacramento

Need to reproject data (transform) before we an do spatial operations.

```{r}
# transform to 3310 to match
gwl_sac_stations_3310 <- st_transform(gwl_sac_stations, st_crs(ces))
st_crs(gwl_sac_stations_3310)
st_crs(gwl_sac_stations_3310) == st_crs(ces)

# get county and crop
gpkg <- here("data/sacramento_gsod.gpkg")
st_layers(gpkg)
sac_co <- st_read(gpkg, layer = "sacramento_isd_cnty") %>% 
  st_transform(3310)
st_crs(sac_co)

# now crop the CES data by the sacramento co polygon
ces_sac <- st_intersection(ces, sac_co) # x, y

mapview(ces_sac)
```

Select points that only fall within certain tract?

```{r}
#tract number: 6077004001
ces_detail <- ces_sac %>% 
  filter(tract == 6077004001)
mapview(ces_detail, col.regions="yellow") + mapview(ces_sac) +
  mapview(gwl_sac_stations, col.regions="orange")

```



```{r}

# load the isd_history file (which ships with {GSODR}):
load(system.file("extdata", "isd_history.rda", package = "GSODR"))

# make spatial
isd_history <- as.data.frame(isd_history) %>% 
  st_as_sf(coords = c("LON", "LAT"), crs = 4326, remove = FALSE)  

# filter to US and CA, many sites out in buoys along coast
isd_history_ca <- dplyr::filter(isd_history, CTRY == "US", STATE == "CA")

# get specific counties
ca_co_select <- ca_co %>% filter(name %in% 
                                   c("Sacramento",
                                     "San Joaquin", 
                                     "Sutter"))

# check CRS is the same
identical(st_crs(ca_co_select)$epsg, st_crs(isd_history)$epsg)


```


```{r}
# view!
library(rnaturalearth)
library(rnaturalearthdata)

# get countries
world <- ne_countries(scale = "medium", returnclass = "sf")

# plot
plot(isd_history$geometry, pch = 16, cex = 0.2, col = "gray50")
plot(world$geometry, add = T, border = "gray10")
plot(ca$geometry, col = "maroon", add = TRUE)
title("GSOD Climate Stations")
```

## SPATIAL JOINS

```{r}

# SPATIAL JOIN: Get CA county POLYGONS that contain ISD points 
# does not bring attributes from points forward
ca_co_isd_poly <- ca_co_select[isd_history_ca, ]

# anti_join: find counties that don't contain ISD points
ca_co_poly_anti <- ca_co[!lengths(st_intersects(ca_co, isd_history_ca)), ]

# plot of all CA counties with and without ISD stations
plot(ca_co$geometry, col = alpha("gray70", 0.3))
#plot(ca_co_select$geometry, col = alpha("gray", 0.3))
plot(ca_co_poly_anti$geometry, col = alpha("purple", 0.3), add = TRUE)
plot(ca_co_isd_poly$geometry, col = alpha("forestgreen", 0.6), add = TRUE)
plot(isd_history_ca$geometry, add = T, pch = 21, bg = "yellow", 
     cex = 0.5, col = "black")
title("Counties with GSOD ISD Stations (green)\n No Stations (purple)")


isd_select <- st_join(isd_history_ca, ca_co_isd_poly[,"countyfp"], left = FALSE)

# how to add attributes

mapview(isd_select, zcol = "name", layer.name = "County") +
  mapview(ca_co_isd_poly, alpha = 0.5, legend = FALSE)

```



# DataRetrieval Package

```{r}
library(dataRetrieval)
# library(nhdplusTools)

all_nldi <- findNLDI(location =c(-121.418, 38.555),
                          nav  = c("DM"), 
                          find = c("nwis", "flowlines", "basin"))

        
                                  
```

