---
title: "rmd_live_code"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Live Code!

This is live! Meaning anytime we save this file, it will appear online in a plain text format, [**here**](https://dl.dropboxusercontent.com/s/fbcybzmhb7phgph/r4wrds_live_rmd.Rmd?dl=0).

Previous versions can be found [here](https://www.github.com/r4wrds/discussion/main/live_code/).


## Installing Packages that aren't on CRAN

```{r, eval=FALSE, echo=TRUE}

# need to install devtools
install.packages("devtools")
library(devtools)
# then packages like urbanmapr
devtools::install_github("UrbanInstitute/urbnmapr")

```


## Load Libraries

Make sure to load libraries. 

```{r, message=FALSE}

# GENERAL PACKAGES
library(here) # path management
library(tidyverse) # data wrangling & viz
library(purrr) # iteration
library(janitor) # name cleaning
library(glue) # pasting stuff together
library(here) # file paths

# SPATIAL PACKAGES
library(sf) # analysis tools
library(mapview)  # interactive maps!
mapviewOptions(fgb = FALSE) # to save interactive maps
# devtools::install_github("UrbanInstitute/urbnmapr")
library(urbnmapr) # county & state boundaries
library(units) # for convert spatial and other units
library(dataRetrieval) # download USGS river data
library(tmap) # mapping
library(tmaptools) 

```

## Import Spatial Data and Project

```{r}

# get CA
ca <- urbnmapr::get_urbn_map(map="states", sf=TRUE) %>% 
  filter(state_abbv == "CA")
# warning is ok!

# get CA county boundary
ca_co <- urbnmapr::get_urbn_map(map = "counties", sf=TRUE) %>% 
  filter(state_abbv == "CA")

st_crs(ca_co)$epsg

# shapefile
sac_co <- sf::st_read(here("data/shp/sac/sac_county.shp"))

# preview
mapview(sac_co)

# check CRS (coord ref system)
st_crs(sac_co)$epsg
st_transform(sac_co, 4326)

# to transform from another spatial datast
sac_co <- st_transform(sac_co, st_crs(ca_co))
st_crs(sac_co)$epsg

# check CRS are the same
st_crs(sac_co) == st_crs(ca_co) 

```


## Visualize: Interactive 

```{r}

sac_box <- st_make_grid(sac_co, n=1)

sac_grid_10 <- st_make_grid(sac_co, n=10, square = FALSE)

mapview(sac_box) + mapview(sac_grid_10) +
  mapview(sac_co, color="orange", lwd=2)


```

## Static Map with Base `plot()`

```{r}
# if we don't specify the geometry, base plot will plot 
# a map for every attribute (columns) in dataset
# plot(ca_co)
plot(ca_co$geometry)

# make it more fancy
# make sure we have all the pieces with a quick test plot
plot(ca$geometry, col = alpha("gray", 0.5), border = "black", lwd=2)

# try sac co by itself
#plot(sac_co$geometry, border = "cyan4", col = "skyblue",alpha=0.4, lwd = 2)

# add layers, need to use add = TRUE
plot(ca_co$geometry, add = TRUE, border = "purple", col = NA)
plot(sac_co$geometry, add = TRUE, border = "cyan4", col = "skyblue",alpha=0.4, lwd = 2)
plot(sac_box, add = TRUE, border = "orange", col = NA, lwd = 1.4)
plot(sac_grid_10, add = TRUE, border = "green", col = NA, lwd = 0.5)

```


## Check Projections

```{r}

# iterate through and check CRS
map(list(ca_co, ca, sac_co), ~st_crs(.x)$epsg)


```

## Iteration and Intersection

```{r}
# read the stations
gw_files <- list.files(path = here("data/gwl/county"),
                       full.names = TRUE, pattern = "*.csv")
gw_files

# read all files into dataframes and combine with purrr
gw_df <- map_df(gw_files, ~read.csv(.x))

table(gw_df$COUNTY_NAME)

# the readr package will also do this same thing by default
# when passed a list of files:
# needs to be csvs with same data types and column number
gw_df <- readr::read_csv(gw_files)

gw_df <- st_as_sf(gw_df, coords = c("LONGITUDE", "LATITUDE"), crs=4326, remove=TRUE) %>% 
  st_transform(., st_crs(ca_co))

class(gw_df)

# preview!
mapview(gw_df, zcol="COUNTY_NAME", layer.name="GW Stations") +
  mapview(sac_co, legend=FALSE)

```

## Spatial Intersection

We can use [] to subset or crop our data spatially.

```{r}
# gw_df[, ] # access dataframe with [row, column]

gw_sac_join1 <- gw_df[sac_co, ]

table(gw_sac_join1$COUNTY_NAME)

# interestingly, the county boundaries don't line up
mapview(gw_sac_join1, zcol="COUNTY_NAME") + mapview(sac_co, color="orange") + mapview(ca_co, col.regions="gray")

```

```{r}

# st_join adds the attributes from the "y" object 
# to the x object
gw_sac_join2 <- st_join(gw_df, sac_co, left=FALSE)


```


## Getting River Data

Use GW Sac as example

```{r}

sac_loi <- gw_sac_join1 %>% filter(STN_ID %in% c("52418", "5605"))
mapview(sac_loi)

# this package pulls flowline/basin/WQ data points
library(dataRetrieval)
st_crs(sac_loi)$epsg

sac_loi <- st_transform(sac_loi, 4326)

# now we can go get flowline data!
us_nwis <- map(sac_loi$geometry,
                ~findNLDI(location = .x,
                          nav  = c("UM"), # UT, DM, DD 
                          find = c("nwis", "flowlines"),
                          distance_km = 100))

names(us_nwis)

us_nwis <- set_names(us_nwis, nm=glue("id_{sac_loi$STN_ID}"))

# more iteration to combine flowlines and WQ sites
us_flowlines <- map_df(us_nwis, ~rbind(.x$UM_flowlines))
us_nwissite <- map_df(us_nwis, ~rbind(.x$UM_nwissite))

mapview(sac_loi, col.region="purple", legend = TRUE, 
        cex=3, layer.name="GW LOI") +
  mapview(us_nwissite, col.regions="orange", 
          legend = TRUE, layer.name="UM NWIS Sites") +
  mapview(us_flowlines, color="steelblue", lwd=2, 
          layer.name="UM Flowlines", legend=FALSE)

```

## Pull River Data 

```{r}
# site: 11446500

# see if there's daily discharge/wtemperature data available ("dv"):
dataRetrieval::whatNWISdata(siteNumber = "11446500",
                            service = "dv", 
                            parameterCd = c("00060", "00010"),
                            statCd = "00003")


## get that data
usgs_Q <- readNWISdv("11446500", 
                parameterCd = "00060", 
                startDate = "2016-10-01") %>% 
  renameNWISColumns()

# plot:
# if weird error, can try dev.off() to reset graphics device in R
library(ggplot2)
(hydro <- ggplot() + 
   geom_line(data = usgs_Q, aes(x = Date, y = Flow, col = site_no),
             size = .5) + 
   #scale_color_brewer(palette = "Set1") +
   #facet_wrap(~site_no, scales = "free_x") + 
   theme_classic() + 
   labs(title="USGS Discharge Data",
        x="", y="Discharge (cfs)") +
   theme(legend.position = "none"))

```

