---
title: "rmd_live_code"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Live Code!

This is live! **Meaning** anytime we save this file, it will appear online in a plain text format, [**here**](https://dl.dropboxusercontent.com/s/fbcybzmhb7phgph/r4wrds_live_rmd.Rmd?dl=0).

Previous versions can be found [here](https://github.com/r4wrds/discussion/tree/main/live_code).

## Load Libraries We Need

```{r libraries, echo=TRUE}

```

# Let's start spatial R!

Documentation for the package sf is here: https://r-spatial.github.io/sf/index.html


# read in shapefiles from sacramento county


```{r}
# to insert an r chunk, control alt i

library(sf)
library(tidyverse)

# unzip the shapefile folder
unzip('data/shp/sac_county.zip', exdir = 'data/shp/sac/')

# read in a shapefile
sac <- st_read('data/shp/sac/sac_county.shp')

# a quick way to see what you're working with
plot(sac) # look at the file
st_geometry(sac) %>% plot # look at geometry
head(sac) # look at the data

names(sac)

```


# How do you read in a csv file and make it spatial?


```{r}
# read in groundwater data that is also geolocated
stations_df <- read_csv('data/gwl/stations_sac.csv')
stations_df

print(stations_df, width = Inf)

# to convert the df to a sf object
stations <- st_as_sf(stations_df, # this is the df 
         coords = c("LONGITUDE", "LATITUDE"), #cols for cords
         crs = 4269,
         remove = F) #don't remove lat/long
class(stations)


# what happens if coords is a vector? it also works!
st_as_sf(stations, # this is the df 
         coords = c(stations$LONGITUDE, stations$LATITUDE), #cols for cords
         crs = 4269,
         remove = F) #don't remove lat/long

# seee what this is
plot(st_geometry(stations))

```

I want to work with the polygon and points together. 

```{r}
st_crs(sac)
st_crs(sac)$epsg # a way to get the espg out
st_crs(stations)$epsg

# I want these to agree
st_transform(x = sac, crs = 4269)
sac <- st_transform(x = sac, crs = st_crs(stations))


# read in stations for all of california

all_gw_stations <- read_csv('data/gwl/stations.csv') %>% 
  st_as_sf(coords = c('LATITUDE', "LONGITUDE"))

plot(st_geometry(all_gw_stations))

```

I want just the staitons in Sacramento county. How can I do that? 
- filter from the data frame!
- use the spatial objects!

```{r}
names(all_gw_stations)
unique(all_gw_stations$COUNTY_NAME)
# fiilter the sf dataframe
stations_sac <- all_gw_stations %>% # control shift m to pipe
  filter(COUNTY_NAME == 'Sacramento')


# use the polygon to filter
# this is a good resource for spatial data: https://geocompr.robinlovelace.net/geometry-operations.html
# st_intersection = gives you data where x and y intersect

all_gw_stations <- read_csv('data/gwl/stations.csv') %>% 
  st_as_sf(coords = c('LONGITUDE', 'LATITUDE'), # remember this is the ordr of coordinates!!!
           crs = 4269)

# peak at this
plot(st_geometry(all_gw_stations))

st_crs(sac)$epsg

# transform it
sac_transformed <- st_transform(sac, st_crs(all_gw_stations))

# now we can do manipulations on the geometry 
stations_in_sac <- st_intersection(all_gw_stations, sac_transformed)

# another way to filter your data is by indexing
stations_in_sac2 <- all_gw_stations[sac_transformed,]
plot(st_geometry(stations_in_sac2))


```


Another way to plot the data is with ggplot


```{r}
ggplot() + # start a ggplot object
  geom_sf(data = sac_transformed) + # add polygon
  geom_sf(data = stations_in_sac2, aes(color = WELL_DEPTH)) + # add points
  scale_color_viridis_c(begin = .1, end = .8)


methods(class = 'sfc')# returns functiosn taht work with objectss of class sfc
methods(class = 'sf')
class(sac_transformed)

```


You can use mapview to viz spatial and interact with it 

```{r}
library(mapview)

mapview(sac) +
  mapview(stations_in_sac2, cex = 5)

```









