---
format: 
  html:
    embed-resources: true
    theme: materia
params:
  county: Sacramento
---

# `r paste(params$county, 'Groundwater Levels')`

<br>

```{r data-package-load, echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}

# load libraries
library(tidyverse)
library(here)
library(glue)
library(sf)
library(mapview)
library(maps)
library(DT)
mapviewOptions(fgb = FALSE)

# set defaults
knitr::opts_chunk$set(echo= FALSE, warning = FALSE, message = FALSE, out.width = "100%")

# load our data
gwl <- read_rds(glue("{here::here()}/data/sac_yolo_sj.rds"))

# filter all groundwater level data (already loaded into memory) by 
# the supplied county
d <- filter(gwl, COUNTY_NAME == params$county)

# extract the county spatial file
county_sf <- st_as_sf(map("county", plot = FALSE, fill = TRUE)) %>% 
  filter(ID == paste0("california,", tolower(params$county)))

```


This report shows groundwater levels in `r params$county` county.

Dates range from `r min(d$MSMT_DATE, na.rm = TRUE)` to `r max(d$MSMT_DATE, na.rm = TRUE)`.

Data source: [DWR Periodic Groundwater Level Database](https://data.cnra.ca.gov/dataset/periodic-groundwater-level-measurements).

<br>

## Distribution of measurements over time

50% of measured values occur on or after `r median(d$MSMT_DATE, na.rm = TRUE)`.

```{r hist, echo = TRUE}
d %>% 
  ggplot() +
  geom_histogram(aes(MSMT_DATE)) +
  theme_minimal() +
  labs(title = "", x = "", y = "Count")
```


<br>

## Monitoring sites

```{r map}
# mapview of county outline
county_mv <- mapview(
  county_sf, layer.name = paste(params$county, "county"), 
  lwd = 2, color = "red", alpha.regions = 0
)

# mapview of monitoring points
points_mv <- d %>% 
  group_by(SITE_CODE) %>% 
  slice(1) %>% 
  select(-MSMT_DATE) %>% # remove msmt date b/c its irrelevant
  mapview(layer.name = "Monitoring stations")

county_mv + points_mv
```

<br>

## All groundwater levels

```{r plot, echo = FALSE}
# interactive hydrograph
p <- ggplot(d, aes(MSMT_DATE, WSE, color = SITE_CODE)) + 
  geom_line(alpha = 0.5) +
  guides(color = FALSE)
#plotly::ggplotly(p)
p
```

<br> 

```{r dt, echo = FALSE}
# data table of median groundwater level per site, per year
d %>% 
  select(-c("COUNTY_NAME", "WELL_DEPTH")) %>%
  st_drop_geometry() %>% 
  mutate(YEAR = lubridate::year(MSMT_DATE)) %>% 
  group_by(SITE_CODE, YEAR) %>% 
  summarise(WSE_MEDIAN = median(WSE, na.rm = TRUE)) %>%
  ungroup() %>% 
  DT::datatable(
    extensions = 'Buttons', options = list(
      dom = 'Bfrtip',
      buttons = 
        list('copy', 'print', list(
          extend  = 'collection',
          buttons = c('csv', 'excel', 'pdf'),
          text    = 'Download'
        ))
    )
  )
```


***

Report generated on `r Sys.Date()`.